A multi-scalar model for system of cities

### 2019-09-29 Docs/Communications/2019/CCS2019/Communication/slides.tex
slides_v0

@@

@@

@@

@@

@@

@@

\sframe{Model Rationale}{

@@

@@

@@

@@

@@

@@

@@

@@

%Further work will consist in more targeted simulation experiments, including specific exploration algorithms such as diversity search for model regimes \cite{reuillon2013openmole}, to test the model as a proof-of-concept of models for policies. Such a model can also be calibrated on real city systems and urban form trajectories, to extrapolate coupling parameters that would be difficult to obtain otherwise.

@@

$\rightarrow$

$\rightarrow$

@@

\textbf{Open repository for model and results at}

\texttt{https://github.com/JusteRaimbault/UrbanGrowth}

\textbf{Simulation data at}

\texttt{}% dataverse

@@

>>>>

% Q : integrate macro indics ? bof, not used the more elaborated ones

% striking image intro

% history of Singapore: emerged within a multi-scale context - and then local growth (strongly shaped by its interactions at many scales)

@@

\begin{columns}
	\begin{column}{0.5\linewidth}

		% includegraphics

		% Nedum model ?

		\includegraphics[width=\linewidth]{figures/nedum.png}

		\medskip

		\textit{Land-use transport models}

		\cite{viguie2014downscaling}

	\end{column}
	\begin{column}{0.5\linewidth}

		\begin{center}
	\includegraphics[width=0.85\textwidth]{figures/intro_RBD_lattice.png}
	\end{center}

	%\footnotesize
\textit{Hybrid urban morphogenesis model}

	\cite{raimbault2014hybrid}

 %Raimbault, J., Banos, A., \& Doursat, R. (2014, June). A Hybrid Network/Grid Model of Urban Morphogenesis and Optimization. In 4th International Conference on Complex Systems and Applications (pp. 51-60).

	\end{column}
\end{columns}

@@

\begin{center}
\includegraphics[width=0.55\textwidth]{figures/simpop1.png}
\includegraphics[width=0.42\textwidth]{figures/setup_synth_1_tick100.png}
\end{center}

\medskip

\textit{The series of Simpop models: from Simpop1 \cite{sanders1997simpop} to SimpopNet \cite{schmitt2014modelisation}}

@@

% images scales Denise

\begin{center}
\includegraphics[height=0.65\textheight]{figures/evoltheory_scales}
\end{center}

\textit{Scales of urban systems of systems described by the evolutionary urban theory} \cite{pumain2018evolutionary}

%\textit{Systems of cities as co-evolutive systems in which interactions are crucial}
%\cite{pumain1997pour}
%\cite{pumain2008socio}
%\cite{pumain2018evolutionary}

@@

\justify

$\rightarrow$ Weak inter-scale coupling, such as progressive resolution for land-use model, does not considers emergence and autonomous scales

$\rightarrow$ An integrated model, or strongly coupled, would be \textit{a new model extending the two coupled models in the sense that it includes them in some parameter settings or limit conditions}

\textbf{\textit{Urban multi-scalar complexity must be capture by a strongly coupled model}}

@@

$\rightarrow$ no strongly coupled multi-scalar model in the literature

\medskip

$\rightarrow$ need to however consider ``simple'' models to be able to understand their behavior and extract knowledge from them

\textbf{Research objective: }

\medskip

\textit{Investigate a strong coupling of a simple urban system interaction network at the macroscopic scale with an urban morphogenesis model at the mesoscopic scale}

@@

\sframe{Model rationale}{

\includegraphics{figures/multiscale_morph.pdf}

\sframe{Model rationale}{

@@

\textbf{Model scales and objects: }

\begin{itemize}
	\item Cities as agents at the macroscopic scale
	\item Cities as population density grids at the mesoscopic scale
\end{itemize}

\medskip

\textbf{Processes: }

\begin{itemize}
	\item Interaction flows between cities and endogenous growth
	\item Urban sprawl and aggregation within urban areas
	\item \textbf{Downward feedback} Adaptation of urban form parameters to population and accessibility growth (``policy response'')
	\item \textbf{Upward feedback} Adaptation of city interaction range depending on internal performance (congested flows)
\end{itemize}

@@

At each time step:

\medskip

\begin{enumerate}
	\item Macroscopic population are evolved with the interaction model
	\item Population and accessibility differences modify the mesoscopic parameter of diffusion (sprawl) and aggregation (metropolization) \textbf{\textit{(downward feedback)}}
	\item Urban forms are evolved for each city at the mesoscopic scale to match population increases
	\item City interaction ranges are modified according to the internal performance of each city (congested flows) \textbf{\textit{(upward feedback)}}
\end{enumerate}

\sframe{Macroscopic model}{

% rq : does the macro mmodel captures many rank clock inversions ? kind of given variations of the rank correlation (cite macrocoevol)

% + detail indicators here

\justify

\textit{System of cities interaction model including network evolution; production of multiple co-evolution regimes and calibration for France.}

\cite{raimbault2018indirect}

\medskip

\begin{center}
\includegraphics[width=0.72\textwidth]{figures/macrocoevol_en.png}
\end{center}

%\nocite{raimbault2019modeling}

%\bigskip
%\tiny

%Raimbault, J. (2018). Indirect evidence of network effects in a system of cities. Environment and Planning B: Urban Analytics and City Science, 2399808318774335.

%\smallskip

%Raimbault, J. (2019). Modeling the co-evolution of cities and networks. In Niel, Z., Rozenblat, C., eds. \textit{Handbook of Cities and Networks}, Edwar Elgar Publishing, \textit{in press}.

\footnotesize
\textit{Indicators at the macroscopic scale: distributions of population, accessibilities, centralities (summarized by average, hierarchy, entropy)}

\sframe{Mesoscopic model}{

% + detail urban form indicators

\textit{Aggregation-diffusion model at the mescoscopic scale, covering most existing urban forms in Europe} \cite{raimbault2018calibration}

\medskip

\begin{center}
\includegraphics[height=0.6\textheight]{figures/density_Fig2}
\end{center}

%\footnotesize\textit{Examples of generated territorial shapes}

\footnotesize 
\textit{Indicators at the mesoscopic scale: urban form captured by Moran index, average distance, hierarchy, entropy}

\sframe{Model parameters}{

		\begin{tabular}{|c|c|c|c|}
	\hline %\tiny
		Type & Parameter & Process & Range \\\hline
		\multirow{4}{*}{Macro} & $g_i \textrm{ = } g_0$ & Endogenous growth & $\left[0 ; 0.05 \right]$ \\
		& $w_i \textrm{ = } w_G$ & Interactions weight & $\left[0 ; 0.01 \right]$  \\
		& $\gamma_i \textrm{ = } \gamma_G$ & Interactions hierarchy & $\left[0 ; 5 \right]$  \\
		& $d_i$ & Interactions decay & $\left[0 ; 1000 \right]$  \\ \hline
		\multirow{4}{*}{Meso} & $\alpha_i$ & Aggregation & $\left[0 ; 5 \right]$  \\
		& $\beta_i$ & Diffusion & $\left[0 ; 0.1 \right]$  \\
		& $t_m$ & Urban growth speed & $\{1 ; 20 \}$  \\
		& $n_d$ & Diffusion steps &  $\left[1 ; 5 \right]$ \\ \hline
		\multirow{4}{*}{Multiscale} & $\delta\alpha$ & Metropolization (downward) & $\left[-0.2 ; 0.2 \right]$  \\
		& $\delta\beta$ & Sprawl (downward) & $\left[-0.2 ; 0.2 \right]$ \\
		& $\delta d$ & Efficiency (upward) &$\left[-0.2 ; 0.2 \right]$ \\
		& $\lambda$ & Congestion cost (upward) & $\left[0 ; 10 \right]$ \\\hline
	\end{tabular}

@@

%\footnotesize
\justify

\textit{Performance constraints: simulate $N$ mesoscopic morphogenesis models in parallel (macroscopic interactions are efficient as based on matrices)}

\medskip

$\rightarrow$ model implemented in \texttt{scala} and integrated within a broader library (including implementations of \cite{raimbault2018indirect} \cite{raimbault2018calibration} \cite{favaro2011gibrat} \cite{cottineau2015modular})

\textit{Large number of parameters and output indicators}

\medskip

$\rightarrow$ integration into the OpenMOLE model exploration open source software \cite{reuillon2013openmole}

\begin{center}
\includegraphics[width=0.55\textwidth]{figures/openmole.png}
\end{center}

(X year of computation for the results presented here)

@@

Model applied on synthetic systems of cities:

\begin{itemize}
	\item random positions and rank-size hierarchy ($\alpha \textrm{ = } 1.0$ and $P_0 \textrm{ = } 100,000$)
	\item countrywide urban system scale: 500km and 20 cities
	\item initial population grids as monocentric (grid of size 50 and center cell density 1000 units)
	\item simulated for 20 macroscopic time steps (order of magnitude of half a century)
\end{itemize}

@@

\textit{For all indicators, median sharpe ratios computing for a parameter point across repetitions are all larger than 1.6}
%(min around 1 for 3 indicators, above 5 for 2 and around 0.5 for 2)}

\medskip

\includegraphics[width=0.8\textwidth]{figures/allindics_hist.png}

\sframe{Model exploration}{

\begin{center}
\includegraphics[width=\textwidth]{figures/onefactor_allindics_mesoMacroDecayUpdateMax_errorbars.png}
\end{center}

\footnotesize
\textit{U-shape behavior of both macroscopic and mesoscopic indicators as a function of $\delta d$}

\sframe{Model exploration}{

% what is interesting: that macro -> meso -> macro has indeed a role (or the other way around)

\begin{center}
\includegraphics[height=0.8\textheight]{figures/deltaHierarchymacroPopulations-mesoTimeSteps_colorMacroInteractionDecay_facetmesoMacroDecayUpdateMax-macroMesoBetaUpdateMax.png}
\end{center}

\footnotesize
\textit{Non-trivial influence of coupled feedbacks on the different scales}

@@

\begin{center}
\includegraphics[height=0.7\textheight]{figures/deltamesoMorans-macroMesoAlphaUpdateMax_colormacroMesoBetaUpdateMax_facetmesoMacroCongestionCost-mesoMacroDecayUpdateMax_mesoBeta0_11.png}
\end{center}

\textit{Mesoscopic centralization appears at a $\delta \beta$ critical value for low $\delta \alpha$; influenced by upward feedback}

@@

\begin{center}
\includegraphics[height=0.7\textheight]{figures/deltamesoSlopes-macroMesoAlphaUpdateMax_colormacroMesoBetaUpdateMax_facetmesoMacroCongestionCost-mesoMacroDecayUpdateMax_mesoBeta0_11.png}
\end{center}

\textit{Mesoscopic hierarchy has a U-shape of $\delta \beta$ in negative $\delta \alpha$, but has a plateau for positive values} 
%$\rightarrow$ sprawl is needed in a sense

% TODO 
%\sframe{Calibration}{
%}

@@

%Further work will consist in more targeted simulation experiments, including specific exploration algorithms such as diversity search for model regimes \cite{reuillon2013openmole}, to test the model as a proof-of-concept of models for policies. Such a model can also be calibrated on real city systems and urban form trajectories, to extrapolate coupling parameters that would be difficult to obtain otherwise. 
\justify

\textbf{Theoretical and practical implications}

\medskip

$\rightarrow$ model effectively captures an interaction between downward and upward feedback: weak emergence \cite{bedau2002downward}

\medskip

$\rightarrow$ coupling ``simple three parameters models'' yield a complicated and complex simulation model: necessity of complexity and simulation models to understand urban complexity? 
% 'suggest in a way failure of reductionist empistemology (in the sense of a la Barthelemy) - at least for grabbing the multi-scalar nature of systems' (seems a tautology though

\medskip

$\rightarrow$ progressive integration towards models for policy?

\textbf{Developments}

$\rightarrow$ diversity search algorithm to find e.g. regimes with the strongest effect of feedback

\medskip

$\rightarrow$ parametrization on real systems; possibly calibration

(see \cite{raimbault2019ilus} forthcoming presentation at ILUS)

@@

$\rightarrow$ A first step towards \textit{strongly} multi-scalar models to capture urban complexity towards policy models

\medskip

$\rightarrow$ Towards integrative models and theories for urban systems \cite{raimbault2019methods}

@@

\footnotesize

\textbf{Open repositories for}

\begin{itemize}
 \item Model: \texttt{https://github.com/JusteRaimbault/UrbanGrowth-model}
 \item Project: \texttt{https://github.com/JusteRaimbault/UrbanGrowth}
 \item Simulation data: \texttt{}% dataverse
\end{itemize}

@@

\sframe{Mesoscopic model}{

$\rightarrow$ Grid world with cell populations $\left(P_i\left(t\right)\right)_{1\leq i\leq N^2}$.

$\rightarrow$ At each time step:

\begin{enumerate}
\item Population growth with exogenous rate $N_G$, attributed independently to a cell following a preferential attachment of strength $\alpha$
%\begin{equation}
%\Pb{P_i(t+1)=P_i(t)+1|P(t+1)=P(t)+1}=\frac{(P_i(t)/P(t))^{\alpha}}{\sum(P_i(t)/P(t))^{\alpha}}
%\end{equation}
%The attribution being uniformly drawn if all population are equal to 0.
\item Population is diffused $n_d$ times with strength $\beta$
\end{enumerate}

$\rightarrow$ Stopping criterion: fixed maximal population $P_m$.

%To avoid bord effects such as reflecting diffusion waves, border cells diffuse their due proportion outside of the world, implying that the total population at time $t$ is strictly smaller than $N_G\cdot t$.

$\rightarrow$ Output measured by morphological indicators: Moran index, average distance, rank-size hierarchy, entropy.

\textit{Macroscopic interactions}

\medskip

\begin{equation}
		P_i\left( t \textrm{ + } 1 \right) \textrm{ = } P_i\left(t\right) \left(1 \textrm{ + } \Delta t \cdot \left(g_i \textrm{ + } \frac{w_i}{N} \cdot \sum_j \frac{V_{ij}}{<V_{ij}>} \right) \right)
	\end{equation}

where the gravity interaction potential is given by

\begin{equation}
		V_{ij} \textrm{ = } \left(\frac{P_i P_j}{(\sum_k p_k)^2}\right)^{\gamma_G} \cdot \exp \left(- \frac{d_{ij}}{d_i} \right)
\end{equation}

\medskip

Accessibility given by

\begin{equation}
	Z_i \textrm{ = } \sum_j \frac{P_j}{\sum_k P_k} \cdot \exp( - d_{ij} / d_i)
\end{equation}

\begin{itemize}
	\item Mesoscopic growth rate
			    $N_G^{\left(i\right)} \left(t \textrm{ + } 1\right) = \Delta P_i / t_m$
	\item Sprawl parameter (population pressure)
		\begin{equation}
			\beta_i \left(t \textrm{ + } 1\right) \textrm{ = } \beta_i \left(t\right) \cdot \left(1 \textrm{ + } \delta\beta \cdot \frac{\Delta P_i (t)}{\max_k  \Delta P_k \left(t\right)}\right)
		\end{equation}
	\item Aggregation parameter (metropolization process)
	    \begin{equation}
			\alpha_i \left(t \textrm{ + } 1\right) \textrm{ = } \alpha_i \left(t\right) \cdot \left(1 \textrm{ + } \delta\alpha \cdot \frac{\Delta Z_i (t)}{\max_k  \Delta Z_k \left(t\right)}\right)
		\end{equation}

\end{itemize}

Congested flows within the mesoscopic zones:

\begin{equation}
		U_i \textrm{ = } \sum_{kl} \left( \frac{P_k P_l}{P^2} \cdot \frac{1}{d_{kl}} - \lambda \left(\frac{P_k P_l}{P^2} \cdot \frac{1}{d_{kl}}\right)^2 \right)
\end{equation}

Update the macroscopic interaction distance accordingly

\begin{equation}
	d_i \left(t \textrm{ + } 1\right) \textrm{ = } d_i \left(t\right) \left( 1 \textrm{ + } \delta d \cdot \frac{U_i}{\max_k \left|U_k\right|} \right)
\end{equation}

\sframe{Morphological indicators}{

\begin{enumerate}
\item Rank-size slope $\gamma$, given by $\ln \left( P_{\tilde{i}}/P_0\right) \sim k + \gamma\cdot \ln \left(\tilde{i}/i_0\right)$ where $\tilde{i}$ are the indexes of the distribution sorted in decreasing order.
\item Entropy of the distribution:
\begin{equation}
\mathcal{E} = \sum_{i=1}^{M}\frac{P_i}{P}\cdot \ln{\frac{P_i}{P}}
\end{equation}
$\mathcal{E}=0$ means that all the population is in one cell whereas $\mathcal{E}=0$ means that the population is uniformly distributed.
\item Spatial-autocorrelation given by Moran index, with simple spatial weights given by $w_{ij} = 1/d_{ij}$
\[
I = M \cdot \frac{\sum_{i\neq j} w_{ij} \left(P_i - \bar{P}\right)\cdot\left(P_j - \bar{P}\right)}{\sum_{i\neq j} w_{ij} \sum_{i}{\left( P_i - \bar{P}\right)}^2}
\]
\item Mean distance between individuals
\[
\bar{d} = \frac{1}{d_M}\cdot \sum_{i<j} \frac{P_i P_j}{P^2} \cdot d_{ij}
\]
where $d_M$ is a normalisation constant
\end{enumerate}
